---
title: "SARS-CoV-2 QC report"
nav: "Quality control"
author: "BIMSB Bioinformatics Platform"
date: '`r format(as.POSIXct(if ("" != Sys.getenv("SOURCE_DATE_EPOCH")) { as.numeric(Sys.getenv("SOURCE_DATE_EPOCH")) } else { Sys.time() }, origin="1970-01-01"), "%Y-%m-%d %H:%M:%S")`'
params:
  sample_name: ""
  coverage_file: ""
  multiqc_report: ""
  logo: ""
  coverage_table_outfile: ""
---

<style>
.dropdown-menu {
  max-height: 200px;
  overflow-y: scroll;}
</style>

<div id="logo" align="top">
`r knitr::include_graphics(params$logo)`
</div>


<sample>`r params$sample_name`</sample>

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(dplyr)
library(DT)
library(magrittr)
library(base64url)
library(stringr)
library(data.table)
```

This report provides an overview over general quality control measures of the 
sample and provides the summarized data as downloadable tables. 
For detailed information about the methods please consult the
[documentation](http://bioinformatics.mdc-berlin.de/pigx_docs/pigx-sars-cov-2.html#output-description).

```{r reading_data, include = FALSE}
# FIXME Rename these col names sensibly upstream
coverage_df <- read.table(
  params$coverage_file,
  sep = "\t",
  header = TRUE,
  # this is necessary, else we have "." instead of spaces in colnames.
  # not having them is preparation for switching to fread
  check.names = FALSE
  ) %>%
  mutate(across(everything(), ~ na_if(.x, "[]"))) %>%
  mutate(across(everything(), ~ gsub("\\[|\\]", "", .x))) %>%
  mutate(`Number of mutations not covered` = str_replace_all(
    `Number of mutations not covered`, "\\'", ""
  ))
```

# Coverage reports
Here, the information about read coverage of the reference genome as well as 
coverage of the mutation locations that were provided are listed. From those 
values, the overall quality of the sample can be assessed and whether or not its
analysis results should and will be used for further trend-analysis and 
visualization. 

## Mutation locations

In order to assess variant calling based on mutations of interest it is crucial
that its genome locations are covered by read alignment. The table below shows
the proportion of covered mutation locations, how many mutations are tracked in
total, how many of their locations are covered and which locations have no read
coverage at all. 
 
```{r amplicon_coverage, echo = FALSE}
mutation_coverage_cols <- c(
  "Total number of tracked mutations",
  "Total number of mutations covered",
  "Number of mutations not covered"
)

mut_cov_df <- coverage_df %>%
  select(all_of(mutation_coverage_cols)) %>%
  rename(c(
    "Total number of tracked mutations" = mutation_coverage_cols[1],
    "Number of mutations covered" = mutation_coverage_cols[2],
    "Mutation locations (NT) not covered" = mutation_coverage_cols[3]
  ))
  

datatable(
  t(mut_cov_df) %>% set_colnames("Value"),
  options = list(searching = FALSE, paging = FALSE))
```

## Reference Genome

This table shows the total number of raw reads, number of aligned reads, the 
mean coverage depth and percentage of the reference genome covered by 
pre-processed reads.

```{r read_coverage, echo = FALSE}
read_coverage_cols <- c(
  "Total number aligned reads",
  "Percentage ref.genome covered",
  "Mean depth ref.genome coverage"
)

read_cov_df <- coverage_df %>%
  select(all_of(read_coverage_cols)) %>%
  mutate(
    `Percentage ref.genome covered` = paste(round(as.integer(
      `Percentage ref.genome covered`
      ), 1), "%"),
    `Mean depth ref.genome coverage` = round(as.integer(
      `Mean depth ref.genome coverage`
      ))
    ) %>%
  rename(c(
    "Number of aligned reads" = read_coverage_cols[1],
    "Total ref. genome coverage" = read_coverage_cols[2],
    "Mean read depth" = read_coverage_cols[3])
  )

datatable(t(read_cov_df) %>% set_colnames("Value"))
```

# Read processing

To improve alignment rates and mutation calling the pipeline contains trimming
steps for primer, read-quality and adapter trimming. Parameter for read
processing can be provided by 
[the pipeline’s settings file](link/to/etc/settings_file). For detailed method 
information please
consult the 
[documentation](http://bioinformatics.mdc-berlin.de/pigx_docs/pigx-sars-cov-2.html#introduction).

## MultiQC 

In order to be able to assess and compare the read quality after each step
multiple quality reports are generated which are presented below in a 
summarized format:

Download: [MultiQC report of raw and processed data](`r params$multiqc_report`)

## Trimming

To improve alignment rates and mutation calling, proper trimming of reads should
be considered e.g. based on the provided
fastqc-reports.

# Download Stats csv

```{r data_download, include = FALSE}
write.csv(coverage_df, params$coverage_table_outfile)

coverage <- readLines(params$coverage_file) %>%
  paste0(collapse = "\n") %>%
  base64_urlencode()
```

Coverage statistics:  
[Download Coverage_stats.csv](`r sprintf('data:text/csv;base64,%s', coverage)`)
